---
title: "TS_Project_HTS"
author: "Chinta"
date: "2023-03-03"
output:
  html_document:
    df_print: paged
---
## 0. Import dependencies.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(fpp)
library(fpp2)
library(fpp3)
library(tseries)
library(hts)
library(xts)
library(lubridate)
library(tsibble)
library(tibble)
library(forecast)
library(feasts)
library(fable)
library(knitr)
```

## 1. Load the chickenpox data.

Load the data, get the region-county data, join the chickenpox cases with region-county data, and make it a monthly structure.

```{r echo=FALSE}
hch <- read.csv("hungary_chickenpox.csv")  #read the data

hch_r_c <- read.csv("hungary_rgn_county.csv") #get the region-counties info

names(hch_r_c) <- c("region", "rgn","county") #rename the columns

hch.df <- data.frame(hch) #Change to data frame

hch.df <- hch.df  %>%  
  pivot_longer (cols =! Date,names_to = "county", 
                values_to ="cases_count") #Unpivot the data

hch.df <- hch.df %>% left_join(hch_r_c) #Join the cases data with region-counties data

hch.df <- hch.df %>% select ( Date, region, rgn, county, cases_count) #arrange the columns

hch.df$Date <- as.Date(hch.df$Date, format="%d/%m/%Y") #format the date using as.Date

#hch.df <- hch.df %>% unite ("rgn", rgn:county, sep="") #combine the region and county columns

hch.df <- hch.df %>% select ( Date, region,  county, cases_count) #arrange the columns

hch.df$Date <- yearmonth(hch.df$Date) #format the Date columns to change the date from weekly to monthly

hch.df <- hch.df %>% group_by(Date, region, county) %>% summarise (cases_count = sum(cases_count)) 
```

## 2. Create a training and a test tsibble.

Creating the training and test tsibble. Dates less than December 2013 are part of training dataset, while above 12/2013 is test dataset. Additionally, use aggregate_key function to create country-level cases count.

```{r echo=FALSE}
hch.tsb <- as_tsibble(hch.df, key=c(region, county), index = Date) #create a tsibble object

#Check for gaps in the data
scan_gaps(hch.tsb) %>% 
  count(Date) 

hch.tsb_agg <- tsibble(hch.df, key = c(region, county), index = Date) %>% 
  aggregate_key(region/county, cases_count = sum(cases_count))

hch.tsb_agg_train <- hch.tsb_agg %>% filter(year(Date) <= 2013)
hch.tsb_agg_test <- hch.tsb_agg %>% filter(year(Date) > 2013)

```

## 3. Plot aggregate (country level) and Central Hungary training data.

Plot the aggregate and Central Hungary training data to understand data. We see that Central Hungary has pretty heavy influence on the country-level data.

```{r echo=FALSE}
#The variable that we'd like to estimate is the number of cases represented by the 'cases_count' variable. The plot reveals that weak trends and high seasonality are apparent.

hch.tsb_agg_train %>% 
  filter(is_aggregated(region)) %>% 
  autoplot(cases_count) #plot the output

hch.tsb_agg_train %>% 
  filter(region == 'Central Hungary') %>% 
  autoplot(cases_count) #plot the output
```

## 3. Model ETS on cases_count with additive trend and reconcile using bottom_up, top_down, and minimum trace methodlogy.



```{r echo=FALSE}

fit <- hch.tsb_agg_train %>% 
  model(base = ETS(cases_count ~ trend("A"))) %>%  
  #model(arima = ARIMA(cases_count))%>%
  reconcile(bottom_up = bottom_up(base),
            top_down = top_down(base),
            min_trace = min_trace(base, "mint_shrink"))

fit_all <- hch.tsb_agg_train %>% 
  model(ets = ETS(cases_count ~ trend("A")),
        arima = ARIMA(cases_count)) %>%  
  reconcile(bu_ets = bottom_up(ets),
            td_ets = top_down(ets),
            min_trace_ets = min_trace(ets, "mint_shrink"),
            bu_arima = bottom_up(arima),
            td_arima = top_down(arima),
            min_trace_arima = min_trace(arima, "mint_shrink")) 


```

```{r echo=FALSE}
fit %>% 
  select(region, county, base) %>%
  filter(region=="Central Hungary") %>% 
  coef()

fc_f <- fit %>% 
  forecast(new_data = hch.tsb_agg_test) 

fc_all <- fit_all %>% 
  forecast(new_data = hch.tsb_agg_test) 

```

Plot the forecast at the country-level and all the regions

```{r echo=FALSE}
autoplot(fc_f, hch.tsb_agg %>% ungroup() %>% filter(year(Date) >2012), level = NULL) + 
  facet_wrap(~region, scales = "free_y") + 
  scale_y_continuous(labels = scales::comma_format()) + 
  labs(color = "", x = "Date" ,y = "Number of Fares") + 
  theme(
    legend.position = 'bottom',
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
  )

```

```{r echo=FALSE}
fc_f %>% 
  filter(is_aggregated(region)) %>% 
  autoplot(hch.tsb_agg %>% filter(year(Date) >2012), level = NULL )

```

```{r echo=FALSE}
fc_all %>% 
  filter(is_aggregated(region)) %>% 
  autoplot(hch.tsb_agg %>% filter(year(Date) >2012), level = NULL )

```